### 2019-04-02 thesis-template/msc-thesis.tex
add

\section{Objectives}
The problem setting encomprises the visualization of tensor fields, i.e. mostly 2D/3D-grids consisting of mainly second-order tensors (cnf. Cauchy-Stress Tensor). The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies e.g. clustering, structuring or projection techniques. The objective of this work is to develop a simple light propagation scheme to generate a tensor-induced flow map. The tensors are decomposed via singular value decomposition to obtain a real eigenbasis as transmission profile. Another aim is to implement a vector-field visualization borrowed FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian Coherent Structures). These structures can be identified as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of differing stress distributions. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions. The real part of the eigenvalues can be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the stress tensor just the principal normal stresses. Thus, we can even define the full set of stresses with the deformation glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order). In fact, for normal square (symmetric and antisymmetric matrices) the singular values are related to the eigenvalues by square root ($s_i = \sqrt{|\lambda_i|}$).

@@

 Now for the rest of this work, we will give a quick overview about the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like Tensors and PCA in the context of singular value decomposition. Also we will discuss the most relevant work in the domain of tensor field visualization. In chap. \ref{chap:main} we will propose the main contribution of this work and document the experimental evaluation in chap. \ref{chap:eval}. The Summary, Conclusion and Outlook will be given in chap. \ref{chap:concl}.\\

@@

In this chapter, we will give the prerequisites to understand the contributions of this work and we will start right away with tensors in general.\\

@@

\section{Principal Component Analysis}
\label{sec:pca}
Note that we talk about a matrix as a second order tensor interchangably for the sake of simplicity. For any state of stress in equilibrium it is possible to find $n$ independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by ellipsoid glyphs.
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/HAT.png}
  }
  \caption{Principal Component Analysis}
  \label{HAT}
\end{figure}

Now, principal component analysis can be done via eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for non-(anti-)symmetric matrices. For (anti-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of tensor. In addition, we also calculate the eigenvalues to exploit the sign of the eigenvalues ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which allows us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is used as a transmission profile for the propagation scheme in symbolic form in a subequent step.
\vskip 5pt
\section{Related Work}
\label{sec:related}

>>>>

\section{Objectives}
The problem setting encomprises the visualization of tensor fields, i.e. mostly 2D/3D-grids consisting of mainly second-order tensors (cnf. Cauchy-Stress Tensor). The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies e.g. clustering, structuring or projection techniques. The objective of this work is to develop a simple light propagation scheme to generate a tensor-induced flow map. The tensors are decomposed via singular value decomposition to obtain a real eigenbasis as transmission profile. Another aim is to implement a vector-field visualization borrowed FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian Coherent Structures). These structures can be identified as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of differing stress distributions. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions. The real part of the eigenvalues can be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the deformation glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order).

@@

For the rest of this work, we will give a quick overview about the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like Tensors and PCA in the context of singular value decomposition. Also we will discuss the most relevant work in the domain of tensor field visualization. In chap. \ref{chap:main} we will propose the main contribution of this work and document the experimental evaluation in chap. \ref{chap:eval}. The Summary, Conclusion and Outlook will be given in chap. \ref{chap:concl}.\\

@@

In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and will capture Tensors, PCA and Related Work in that sense.\\

@@

\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix as a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by ellipsoid glyphs.
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/HAT.png}
  }
  \caption{Principal Component Analysis}
  \label{HAT}
\end{figure}

Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for non-(anti-)symmetric matrices. For (anti-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of tensor as suggested by \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the eigenvalues ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which allows us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step.
\vskip 5pt
\section{Related Work}
\label{sec:related}
